---
title: "Lab4"
author: "Andrew Edelblum"
date: "2/4/2019"
output: html_document
---

```{r setup}
library(rio)
library(janitor)
library(ggplot2)
theme_set(theme_minimal())

honey <- import("./d.csv", setclass = "tibble")
```

# 1. Visualize the total production of honey across years by state.

```{r heatmap1}
ggplot(honey, aes(x = year, y = state, fill = totalprod)) +
  geom_tile() +
  scale_fill_distiller(palette = "Blues") +
  labs(x = "Year", y = "State", title = "Honey Production by Year by State", fill = "Total Honey Produced") +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 16, face = "bold"),
        legend.background = element_rect(size = .5))
```

## Use color to highlight the west coast (Washington, Oregon, and California).

```{r highlights}
# Vector with all unique states in honey dataframe
states <- unique(honey$state)

# Assigns different color to the west coast states
label_color <- ifelse(states == "OR" | states == "WA" | states == "CA", "#C55644", "gray30")

# Assigns bold type to the west coast states
label_face <- ifelse(states == "OR" | states == "WA" | states == "CA", "bold", "plain")
```

```{r heatmap_revised}
(p <- ggplot(honey, aes(x = year, y = state, fill = totalprod)) +
  geom_tile() +
  scale_fill_continuous(type = "viridis") +
  labs(x = "Year", y = "State") +
  theme(axis.text.y = element_text(color = label_color, face = label_face),
        plot.title = element_text(size = 16, face = "bold", hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5),
        legend.title = element_text(size = 16, face = "bold"),
        legend.background = element_rect(size = .5)))
```

# 2. Reproduce the plot according three different kinds of color blindness, as well as a desaturated version.

```{r color_blind}
library(colorblindr)

colorblindr::cvd_grid(p)
```

# 3. Reproduce the plot using a color blind safe pallette.

```{r}
ggplot(honey, aes(x = year, y = state, fill = totalprod)) +
  geom_tile() +
  scale_fill_viridis_c(option = "magma") +
  labs(x = "Year", y = "State") +
  theme(axis.text.y = element_text(color = label_color, face = label_face))
```

# 4. Download the file here denoting the region and division of each state.

```{r}
region <- import("./region.csv", setclass = "tibble") %>% 
  clean_names()


```

## Join the file with your honey file.


